### 2016-09-28 text/ms.Rmd
Edits dealing with reviewer comments

```{r, echo=FALSE}
# Values to use in the paper:
load("values.rda") # generated by `../analysis/9-output-values.R`
if (mean_sim_basic$mach_mare_range == "0.32--0.32") mean_sim_basic$mach_mare_range <- 0.32
# Includes: mean_sim, mean_ram, slope_sim, auc_sim, ram_stocks_n

@@

\noindent where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus
one year, $B_t$ represents biomass at time $t$, $r$ represents intrinsic
population growth rate, $B_0$ represents unfished biomass or carrying capacity
$K$, and $C$ represents catch. The fourth model is an empirically derived model
based on the RAM Legacy Stock Assessment Database [@ricard2012]. @rosenberg2014
provide a full background on these four methods and code to fit all the models
is available in an accompanying package `datalimited` for the statistical
software R [@r2015] (<https://github.com/datalimited/datalimited>) (*Private while in
review*). In summary:

@@

We first developed and tested ensemble methods on a fully factorial simulated
dataset of fisheries [@rosenberg2014] with known status. These simulated data
included the following scenarios: three fish life histories: small pelagic,
demersal, and large pelagic; three levels of initial biomass depletion
compared to carrying capacity: biomass at 100%, 70%, and 40% of carrying
capacity; and four exploitation settings: (1) a constant exploitation rate,
(2) an exploitation rate coupled with biomass to mimic an open-access
single-species fishery, (3) a scenario where exploitation rate increased
continuously, and (4) a "roller-coaster" scenario where the exploitation
rate increased and then decreased. Process noise (recruitment variability;
i.e., unexplained variability in population dynamics) was introduced to the
models at two magnitudes in log space, $N(0, 0.2^2)$ and $N(0, 0.6^2)$,
and was either uncorrelated through time or had first-order autoregressive
correlation of $0.6$. The simulation also included a scenario with $N(0,
0.2^2)$ measurement error around log(catch) and one scenario without
measurement error. @rosenberg2014 ran ten iterations for each combination
of factors adding stochastic draws of recruitment and catch-recording
variability each time to generate a total of 5760 stocks. The simulation
models were built in the `FLR` packages [@kell2007,
<http://flr-project.org>] for the statistical software R. Code to
generate the simulations is available at
<https://github.com/iagomosqueira/stockStatusFAO>.

@@

Superensemble models allow us to incorporate additional covariates and
potentially leverage interactions between these covariates and individual model
predictions. Additional covariates could be, for example, life-history
characteristics, information on exploitation patterns, or statistical
properties of the data. We tested the performance benefits of including one set
of additional covariates: spectral properties of the catch time series.
Spectral analysis decomposes a time series into the frequency domain and
provides a means of describing the cyclical shape of the catch series that is
independent of time series length (except in affecting precision) and
independent of absolute magnitude of catch. We fit spectral models to the
scaled catch time series (catch divided by maximum catch) with the `spec.ar`
function in R and recorded representative short- and long-term spectral
densities at frequencies of 0.20 and 0.05, which correspond to 5- and 20-year
cycles. We include the results of adding these additional covariates in the
supplementary materials.

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database
--- a compilation of stock-assessment output from hundreds of exploited marine
populations around the world. Our analysis of the stock-assessment database was
based on version 2.5. After removing stocks for which at least one of the
individual models described below did not converge (121), this database
included `r ram_stocks_n` stocks. We removed these stocks for all methods ---
both for the individual and superensemble models. An alternative would be to
fit superensemble models to subsets of the individual models that did converge,
but for simplicity we will only use superensemble models fitted to all four
individual models.

In the case of the RAM Legacy Stock Assessment Database, we used superensembles
trained on the entire simulation dataset. However, since mPRM is built on the
same stock-assessment database, we applied three-fold cross-validation to the
data underlying the mPRM model so that the dataset with which mPRM was trained
(for the individual model and superensemble) was separate from the dataset with
which it was tested.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of
bias, precision, accuracy (a combination of bias and precision), or the ability
to correctly rank or correlate across populations [e.g., @walther2005]. Here, we
measure proportional error, defined as $(\hat{\theta} - \theta)/\abs{\theta}$,
where $\hat{\theta}$ and $\theta$ represent estimated and "true" (or
stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median
proportional error to measure bias, median absolute proportional error to
measure accuracy, and Spearman's rank-order correlation between predicted and
"true" values to measure the ability to correctly rank populations. When
testing with the RAM Legacy Stock Assessment database, we treated the
estimates from these data-rich stock assessments as known without error. Thus,
any error in the stock-assessment estimates of the mean or slope of
$B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for
each of the four data-limited models and the ensembles. Code to reproduce our
analysis is available at <https://github.com/datalimited/ensembles> (*Private while in
review*).

@@

Ensemble methods, and in particular the machine learning superensemble models
(random forest and GBM), generally improved estimates of status and trend over
any individual model (Fig. \ref{hexagon}e--h, Fig. \ref{scatter-sim-slope}e--h).
Compared to the individual models, machine learning superensembles decreased
inaccuracy (median absolute proportional error) from
`r mean_sim_basic$ind_mare_range` to `r mean_sim_basic$mach_mare_range`,
increased rank-order correlation from
`r mean_sim_basic$ind_corr_range` to `r mean_sim_basic$mach_corr_range`, and reduced bias
(median proportional error) from `r mean_sim_basic$ind_mre_range` to
`r mean_sim_basic$mach_mre_range` (Fig. \ref{performance}a).
These superensembles also generally had better ability to distinguish if
simulated stocks were above or below $B/B_\mathrm{MSY} = 0.5$ (Fig. \ref{roc-sim}).
Results were similar when predicting trend: compared to individual models,
machine learning superensembles decreased inaccuracy from
`r slope_sim$ind_mare_range` to `r slope_sim$mach_mare_range`,
increased rank-order correlation from
`r slope_sim$ind_corr_range` to `r slope_sim$mach_corr_range`, and reduced
bias from `r slope_sim$ind_mre_range` to `r slope_sim$mach_mre_range` (Fig.
\ref{performance-sim-slope}).
The ensemble models that simply took a mean of the individual models ranked
slightly behind the best individual model for estimating fish stock status
(CMSY; Fig. \ref{performance}a) and had slightly lower correlation but higher
accuracy than the best individual model at predicting the trends of status
(SSCOM; Fig. \ref{performance-sim-slope}).

The superensemble models were able to improve the predictive performance by
harnessing the best properties of individual models, the covariance between
individual models, and interactions with other covariates. For example, SSCOM
had strong predictive ability when it predicted low $B/B_\mathrm{MSY}$ (Fig.
\ref{hexagon}c, Fig. \ref{partial-sim}c) and CMSY predictions were approximately
linearly related to $B/B_\mathrm{MSY}$ within the low and high clusters of
predictions (Fig. \ref{partial-sim}). SSCOM contributed most strongly on its own
to determining trend (Fig. \ref{partial-sim-slope}). Superensembles also
exploited the covariance between individual model predictions. For instance,
both the linear model and GBM ensemble suggest that if mPRM and SSCOM predict
high status, the true status also tends to be high (Figs \ref{lm-coefs},
\ref{partial-2d-sim}l). The addition of spectral density covariates helped the
superensemble models correctly predict higher status values (Fig.
\ref{hexagon-sim-covariates}g, h). The performance of the ensembles was only
marginally improved by including these covariates (Fig.
\ref{performance-with-covariates} vs. Fig. \ref{performance}).

@@

A strength of superensembles is that they can be tailored to predict specific
response variables. For example, we built separate superensemble models of mean
$B/B_\mathrm{MSY}$ and the slope of $B/B_\mathrm{MSY}$. The same set of model
weights or non-linear relationships need not hold across different response
variables. For instance, SSCOM contributed little to the GBM superensemble
estimate of status at higher levels of predicted $B/B_\mathrm{MSY}$ (Fig.
\ref{partial-sim}), but contributed strongly to estimates of trend (Fig.
\ref{partial-sim-slope}). Formally, fitting superensemble models to specific
quantities of interest (such as the slope of $B/B_\mathrm{MSY}$) provides an additional
calibration step [@rykiel1996]. Individual models are typically calibrated
(i.e., parameters are estimated) based on a single response variable; however,
this model might then provide a biased or inaccurate estimate of other
quantities of interest. Ensemble methods provide a final calibration step to a
quantity of interest. This ensemble calibration could include a loss function
tailored to the goals of the model, say placing greater weight on accuracy at
lower rather than higher status levels. Conversely, because superensembles are
tailored to a specific response and loss function, superensembles force a
modeller to choose an operational purpose for their model up front [*sensu*
@dickey-collas2014].

>>>>

```{r, echo=FALSE}
# Values to use in the paper:
load("values.rda") # generated by `../analysis/9-output-values.R`
# if (mean_sim_basic$mach_mare_range == "0.32--0.32") mean_sim_basic$mach_mare_range <- 0.32
# Includes: mean_sim, mean_ram, slope_sim, auc_sim, ram_stocks_n

@@

\noindent where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus
one year, $B_t$ represents biomass at time $t$, $r$ represents intrinsic
population growth rate, $B_0$ represents unfished biomass or carrying capacity
$K$, and $C$ represents catch. The fourth model is an empirically derived model
based on the RAM Legacy Stock Assessment Database [@ricard2012]. @rosenberg2014
provide a full background on these four methods \R{C} 
(<http://www.fao.org/docrep/019/i3491e/i3491e00.htm>, last accessed 2016-09-27) 
and code to fit all the models
is available in an accompanying package `datalimited` for the statistical
software R [@r2015] <https://github.com/datalimited/datalimited> 
(last accessed 2016-09-27). In summary:

@@

We first developed and tested ensemble methods on a fully factorial simulated
dataset of fisheries with known status [@rosenberg2014]. 
Briefly, these simulations were implemented 
with the FLR packages [@kell2007] 
for the statistical software R, and, in particular, the FLBRP package.
The framework takes a series of 
life-history parameters and fishery characteristics
to generate population projections
and resulting catch timeseries.
Life-history values (e.g. mean asymptotic length)
for three fish life histories (small pelagic,
demersal, and large pelagic) 
were translated into a complete set of parameters for
a von Bertalanffy growth model,
a maturity ogive,
natural mortality, 
a selectivity function,
and a Beverton-Holt stock-recruitment function\R{A4}
using the life-history relationships derived in @gislason2008.
Fishing scenarios included three levels of initial biomass depletion
compared to carrying capacity: biomass at 100%, 70%, and 40% of carrying
capacity; and four exploitation patterns: (1) a constant exploitation rate,
(2) an exploitation rate coupled with biomass to mimic an open-access
single-species fishery, (3) a scenario where exploitation rate increased
continuously, and (4) a "roller-coaster" scenario where the exploitation
rate increased and then decreased. Process noise (recruitment variability;
i.e., unexplained variability in population dynamics) was introduced to the
models at two magnitudes in log space, $N(0, 0.2^2)$ and $N(0, 0.6^2)$,
and was either uncorrelated through time or had first-order autoregressive
correlation of $0.6$. The simulation also included a scenario with $N(0,
0.2^2)$ measurement error around log(catch) and one scenario without
measurement error. @rosenberg2014 ran ten iterations for each combination
of factors adding stochastic draws of recruitment and catch-recording
variability each time to generate a total of 5760 stocks. 
Code to generate the simulations is available at
<https://github.com/iagomosqueira/stockStatusFAO> (last accessed 2016-09-27).

@@

Superensemble models allow us to incorporate additional covariates and
potentially leverage interactions between these covariates and individual model
predictions. Additional covariates could be, for example, life-history
characteristics, information on exploitation patterns, or statistical
properties of the data. We tested the performance benefits of including one set
of additional covariates: spectral properties of the catch time series.
Spectral analysis decomposes a time series into the frequency domain and
provides a means of describing the cyclical shape of the catch series that is
independent of time series length (except in affecting precision) and
independent of absolute magnitude of catch. We fit spectral models to the
scaled catch time series (catch divided by maximum catch) with the `spec.ar`
function in R and recorded representative short- and long-term spectral
densities at frequencies of 0.20 and 0.05, which correspond to 5- and 20-year
cycles. For the linear model superensemble, we incorporated the two spectral
covariates (S1, S2) along with all second-order interactions as:\R{D}

\begin{equation}
\hat{\theta_i} = \beta_0 + \beta_1 \hat{b}_{i,1} + ... +
\beta_{1,2} \hat{b}_{i,1} \hat{b}_{i,2} + ... +
\beta_{S1} S1_i + 
\beta_{S2} S2_i + 
\beta_{1,S2} \hat{b}_{i,1} S2_i + ... +
\epsilon_i,
\end{equation}

\noindent
with $\epsilon \sim \mathrm{Normal}(0, \sigma^2)$ and for
simulated fisheries $i$ 1 through $n$.
We include the results of adding these additional covariates in the
supplementary materials.

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database
--- a compilation of stock-assessment output from hundreds of exploited marine
populations around the world. Our analysis of the stock-assessment database was
based on version 2.5. After removing stocks for which at least one of the
individual models described below did not converge (121), this database
included `r ram_stocks_n` stocks. We removed these stocks for all methods ---
both for the individual and superensemble models. An alternative would be to
fit separate superensemble models to subsets of the individual models that did converge,
but for simplicity we will only use superensemble models fitted to all four
individual models.\R{A6}

In the case of the RAM Legacy Stock Assessment Database, we used superensembles
trained on the entire simulation dataset. However, since mPRM is built on the
same stock-assessment database, we applied three-fold cross-validation to the
data underlying the mPRM model so that the dataset with which mPRM was trained
(for the individual model and superensemble) was separate from the dataset with
which it was tested. This meant that, for each iteration of cross validation, 
we split the RAM database into three, fit the mPRM model to two-thirds of
the RAM database, fit a superensemble with this version of mPRM, and then 
tested the performance of the superensemble on the third of the RAM database
we had withheld.\R{A7}

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of
bias, precision, accuracy (a combination of bias and precision), or the ability
to correctly rank or correlate across populations [e.g., @walther2005]. Here, we
measure proportional error, defined as $(\hat{\theta} - \theta)/\abs{\theta}$,
where $\hat{\theta}$ and $\theta$ represent estimated and "true" (or
stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median
proportional error to measure bias, median absolute proportional error to
measure accuracy, and Spearman's rank-order correlation between predicted and
"true" values to measure the ability to correctly rank populations. When
testing with the RAM Legacy Stock Assessment database, we treated the
estimates from these data-rich stock assessments as known without error. Thus,
any error in the stock-assessment estimates of the mean or slope of
$B/B_\mathrm{MSY}$ also contributes to our estimates of prediction error for
each of the four data-limited models and the ensembles. Code to reproduce our
analysis is available at <https://github.com/datalimited/ensembles> (last accessed 2016-09-27).

@@

Ensemble methods, and in particular the machine learning superensemble models\R{A8}
(random forest and GBM), generally improved estimates of status and trend over
any individual model (Fig. \ref{hexagon}e--h, Fig. \ref{scatter-sim-slope}e--h).
Compared to the individual models, machine learning superensembles decreased
inaccuracy (median absolute proportional error) from
`r mean_sim_basic$ind_mare_range` to `r mean_sim_basic$mach_mare_range`,
increased rank-order correlation from
`r mean_sim_basic$ind_corr_range` to `r mean_sim_basic$mach_corr_range`, and reduced bias
(median proportional error) from `r mean_sim_basic$ind_mre_range` to
`r mean_sim_basic$mach_mre_range` (Fig. \ref{performance}a).
These superensembles also generally had better ability to distinguish if
simulated stocks were above or below $B/B_\mathrm{MSY} = 0.5$ (Fig. \ref{roc-sim}).
Results were similar when predicting trend: compared to individual models,
machine learning superensembles decreased inaccuracy from
`r slope_sim$ind_mare_range` to `r slope_sim$mach_mare_range`,
increased rank-order correlation from
`r slope_sim$ind_corr_range` to `r slope_sim$mach_corr_range`, and reduced
bias from `r slope_sim$ind_mre_range` to `r slope_sim$mach_mre_range` (Fig.
\ref{performance-sim-slope}).
The ensemble models that simply took a mean of the individual models ranked
slightly behind the best individual model for estimating fish stock status
(CMSY; Fig. \ref{performance}a) and had slightly lower correlation but higher
accuracy than the best individual model at predicting the trends of status
(SSCOM; Fig. \ref{performance-sim-slope}).

The superensemble models were able to improve the predictive performance by
harnessing the best properties of individual models, the covariance between
individual models, and interactions with other covariates. For example, SSCOM
had strong predictive ability when it predicted low $B/B_\mathrm{MSY}$ (Fig.
\ref{hexagon}c, Fig. \ref{partial-sim}c) and CMSY predictions were approximately
linearly related to $B/B_\mathrm{MSY}$ within the low and high clusters of
predictions (Fig. \ref{partial-sim}). SSCOM contributed most strongly on its own
to determining trend (Fig. \ref{partial-sim-slope}). Superensembles also
exploited the covariance between individual model predictions. For instance,
both the linear model and GBM ensemble suggest that if mPRM and SSCOM predict
high status, the true status also tends to be high (Figs \ref{lm-coefs},
\ref{partial-2d-sim}f). \R{A9}The addition of spectral density covariates helped the
superensemble models correctly predict higher status values (Fig.
\ref{hexagon-sim-covariates}g, h). The performance of the ensembles was only
marginally improved by including these covariates (Fig.
\ref{performance-with-covariates} vs. Fig. \ref{performance}).

@@

A strength of superensembles is that they can be tailored to predict specific
response variables. For example, we built separate superensemble models of mean
$B/B_\mathrm{MSY}$ and the slope of $B/B_\mathrm{MSY}$. The same set of model
weights or non-linear relationships need not hold across different response
variables. For instance, SSCOM contributed little to the GBM superensemble
estimate of status at higher levels of predicted $B/B_\mathrm{MSY}$ (Fig.
\ref{partial-sim}), but contributed strongly to estimates of trend (Fig.
\ref{partial-sim-slope}). Formally, fitting superensemble models to specific
quantities of interest (such as the slope of $B/B_\mathrm{MSY}$) provides an additional
calibration step to a quantity of interest [@rykiel1996].\R{A10}
This ensemble calibration could include a loss function
tailored to the goals of the model, say placing greater weight on accuracy at
lower rather than higher status levels. Conversely, because superensembles are
tailored to a specific response and loss function, superensembles force a
modeller to choose an operational purpose for their model up front [*sensu*
@dickey-collas2014].
